---
title: "subject-17"
author: "Mathilde Vallat"
format:
  html: default
  pdf: default
---
```{r}
#| message: false
here::i_am("eval-grades.Rproj")
library(here)
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
theme_set(theme_bw())

```


## Question 1
Load the data set.

```{r}
grades <- read_delim(here("data", "grades.csv"))

```

## Question 2
Give the number of grades in the data set directly in the text of your quarto file, in a sentence of the form “The data set contains xxx grades.” where xxx is replaced by the number of grades. This must be computed from the data set.

```{r}
total_grades <- nrow(grades)
```

The data set contains 44195 grades.

## Question 3

Use distinct() compute the number of students on the data set and report the value directly in the text as in the previous question.

```{r}
num_students <- grades |> distinct(id) |> summarise(count = n()) |> pull(count)

```

The data set contains 761 students.

## Question 4

```{r}
students <- grades |> distinct(id, group)

students |> slice_head(n = 5) |> knitr::kable()
```

## Question 5

```{r}
grades |> 
  distinct(id, group) |> 
  count(group) |> 
  ggplot(aes(x = group, y = n)) + 
  geom_col() + 
  labs(title = "Number of Students per Group", x = "Group", y = "Student Count")
```

